### 2020-07-03 introduction.tex
editing

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. Climate change has dire consequences on human health and
safety due to extreme weather events and the overall impact on food
production. As a result, urgent measures are necessary to limit greenhouse gas
emissions, namely CO$_2$ emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a large fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location (cite mit).
Employing a diverse mix of these energy sources ensures energy security and
reliability during and after our transition towards a low-carbon future (cite mit).

The world would have to ramp up the current rate of reactor deployments to
displace some of the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy.
A potential solution to the aforementioned issues is
the \gls{MSR} concept, one of six advanced reactor designs
selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
and development.

@@

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
neutronics and thermal-hydraulics are tightly coupled due to the strong
temperature reactivity coefficient. Furthermore, we have to account for the
movement of \glspl{DNP} as
they are now generated directly within the primary coolant loop. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration. Existing reactor system-level software and modeling approaches
for conventional \gls{LWR} analysis contain reactor-specific assumptions in
multiphysics coupling and other areas that render the techniques less suitable
for simulating \glspl{MSR}. Thus, making minor modifications to these codes
without changing the underlying algorithms is not the best approach for
\gls{MSR} safety analysis.

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
\cite{kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion
equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach was featured again in a
later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, in which they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
models showed good agreement in the steady-state neutron flux, temperature,
and \gls{DNP} distributions, and in the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided some quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue such as
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

The preceding discussion highlights the challenges faced in MSR simulation,
and the need for a highly efficient simulation tool that incorporates
implicitly coupled multiphysics with good computational scalability over
multiple processing units. This paper presents the open source MSR simulation
tool, Moltres, as a strong contender to overcoming these challenges for
simulating the MSFR. Moltres is an application code
built in the Multiphysics Object-Oriented Simulation Environment (MOOSE)
parallel finite element framework. Similar to COMSOL and OpenFOAM, it solves
the deterministic multigroup neutron diffusion and thermal-hydraulics
\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
scales well for a large number of processors.

Moltres employs fully implicit coupling between the neutronics and
thermal-hydraulics governing equations, to fully account for the tightly
coupled physics expected in MSRs due to the movement of fuel in the salt. We
implemented flow using the Navier-Stokes equations and zeroth-order
approximation of eddy viscosity for a more accurate representation of the flow
profile compared to imposing a predetermined velocity field.

@@

The main objective of this thesis is to demonstrate Moltres' capabilities in
modeling multiphysics, steady-state and transient behavior of fast-spectrum
\glspl{MSR} through the study of the \gls{MSFR} concept. This is achieved by
first verifying Moltres' neutronics results against Serpent in the context of
the \gls{MSFR}, and then comparing the coupled neutronics/thermal-hydraulics
steady-state and transient accident results of the \gls{MSFR} concept. The
multiphysics results are verified against the results by Fiorina et al.
\cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}.

@@

The outline of this thesis is as follows: Chapter 2 discusses the history and
features of \glspl{MSR}. The \gls{MSFR} concept is covered in greater detail.
Chapter 3 details the simulation codes and the general modeling approach for
generating the results in this thesis. Chapter 4 provides a neutronics
assessment by comparing key neutronics parameters from Moltres' eigenvalue
calculations to Serpent's Monte Carlo calculations. Chapter 5 presents
steady-state results of coupled neutronics/thermal-hydraulics \gls{MSFR}
simulations in Moltres. Chapters 6 to 9 present transient accident simulation
results for unprotected reactivity insertions, unprotected loss of flow,
unprotected pump overspeed, and unprotected loss of heat sink, respectively.
Lastly, Chapter 10 summarizes the key findings of this thesis and posits some
potential avenues for future work.

>>>>

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
atmosphere has been rising steadily throughout modern human history
\cite{us_department_of_commerce_global_nodate}. Climate change has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{noauthor_global_nodate}. As a result,
urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a large fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location
\cite{petti_future_2018}. Employing a diverse mix of these energy sources
ensures energy security and reliability in our transition
towards a low-carbon future \cite{petti_future_2018}.

The world would have to ramp up the current rate of reactor deployments to
displace some of the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy
\cite{massachusetts_institute_of_technology_future_2003}. A potential solution
to the aforementioned issues is the \gls{MSR} concept, one of six advanced
reactor designs selected by the Generation IV International Forum
\cite{gif_technology_2002} for continued research and development.

@@

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
interactions between neutronics and thermal-hydraulics are stronger due
to greater fuel material expansion. Furthermore, \glspl{DNP} in \glspl{MSR}
can flow freely within the primary coolant
loop as opposed to being held in place in a solid fuel matrix. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration.

Most reactor analysis applications are usually reactor-specific by
design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications generally would not
account for \gls{MSR}-specific phenomena and are not suitable for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}, an open source MSR simulation tool. Moltres
is an application built in the \gls{MOOSE} parallel finite element framework.
It solves
the deterministic multigroup neutron diffusion and thermal-hydraulics
\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
scales well for a large number of processors. The underlying \gls{MOOSE}
framework
provides a range of implicit and explicit methods for the coupling between the
neutronics and thermal-hydraulics governing equations.

In the introductory journal article for Moltres, Lindsay et al.
\cite{lindsay_introduction_2018} demonstrated Moltres' capabilities with
2D-axisymmetric and 3D models of the \gls{MSRE}. The results showed good
qualitative agreement with legacy \gls{MSRE} data but there were some minor
quantitative discrepancies due to a number of differences in the legacy model. 
Moltres has undergone further development in the past two years. The authors
from the first paper have since developed various new capabilities in Moltres
such as support for looping delayed neutron precursors back into the core,
and a heat exchanger object to set the heat removal rate.

In addition to those new capabilities, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model. The incompressible Navier-Stokes module provides a more
accurate representation of the flow profile, precursor movement, and heat
transfer as opposed to assuming uniform or function-based velocity fields.

@@

The main objectives of this thesis are to demonstrate and verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because 
there is readily available literature data to verify our results against. The
steady-state and transient results also feature a significant degree of
physics interaction between the neutronics and thermal-hydraulics. This
present work will first present a verification of Moltres' \gls{MSFR}
neutronics results against Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and transient accident
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.

@@

The outline of this thesis is as follows. Chapter 2 discusses the history and
features of \glspl{MSR}, and a literature review of existing \gls{MSR}
simulation tools. The chapter also covers the \gls{MSFR} concept in greater
detail. Chapter 3 details the simulation codes and the general modeling
approach for generating the results in this thesis. Chapter 4 provides a
neutronics assessment by comparing key neutronics parameters from Moltres'
eigenvalue calculations to Serpent's Monte Carlo calculations. Chapter 5
presents steady-state results of coupled neutronics/thermal-hydraulics
\gls{MSFR} simulations in Moltres. Chapter 6 presents transient accident
simulation results for unprotected reactivity insertions, unprotected loss of
heat sink, unprotected loss of flow, and unprotected pump overspeed,
respectively. Lastly, Chapter 7 summarizes the key findings in this thesis
and posits some potential avenues for future work.

### 2020-07-03 model.tex
editing

The main simulation tool used in this work for studying the \gls{MSFR} is
Moltres. While we used Moltres for the steady-state and transient analysis, it
requires a dedicated neutron transport solver such as Serpent 2 to generate
neutron energy group constant data for the multigroup neutron diffusion
calculations in Moltres. This chapter provides brief introductions of the
Serpent and Moltres codes, and the general modeling approach for the
multiphysics simulations in Moltres.

@@

Serpent 2 \cite{leppanen_serpent_2014} is a continuous-energy Monte Carlo
reactor physics code under
active development at the VTT Technical Research Centre of Finland. It was
created in 2004 out of the need for better general-purpose neutron transport
codes for group constant generation in lattice geometries, and has since grown
to support more general capabilities, for an active user base of more than 500
people worldwide. Serpent is highly parallelizable due to its support for
both MPI and OpenMP parallel programming APIs, and highly validated and
verified against experimental data and other well-established codes.

In Serpent, each neutron is tracked through a combination of ray-tracing-based
surface tracking and rejection sampling-based delta-tracking. Users may define
the number of neutrons and the number of active and inactive cycles for each
simulation. Inactive cycles are required for fission source distribution
convergence, before interactions are tallied in the active cycles.
Interaction types and locations are
determined stochastically based on neutron interaction data from established
nuclear data libraries (e.g. ENDF, JEFF). These nuclear data libraries provide
continuous-energy cross section data at discrete temperatures. For other
temperatures, Serpent has a built-in Doppler-broadening preprocessor that
extrapolates the relevant cross section data from a lower temperature.

In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
neutronics benchmark geometry published in several previous papers involving
the \gls{MSFR}, as shown in Fig. \ref{fig:msfrgeom}. The relevant group
constant data are collapsed into six neutron energy groups, and calculated for
temperature values of 800 K to 1300 K at 50 K intervals.

The group constants relevant for neutronics calculations in Moltres are the
macroscopic fission, absorption, and scattering neutron cross sections,
neutron diffusion coefficients, average fission energies, average neutron
yields, inverse neutron speeds, flux spectra, \gls{DNP} decay constants, and
effective delayed neutron fractions. These group constants are extracted from
the Serpent output files using a Python script available from the Github
repository that holds the Moltres source code. The script rewrites the group
constants into a Moltres-compatible format.

\section{Moltres}

Moltres is a \gls{MOOSE}-based application for coupled
neutronics/thermal-hydraulics simulations of \glspl{MSR}. \gls{MOOSE}
\cite{gaston_physics-based_2015} is a highly parallelizable,
finite element framework developed at \gls{INL} for simplifying the process of
creating fully-coupled, non-linear, multiphysics solvers. The framework
provides a user-friendly interface for this task through object-oriented
programming in C++. All aspects of a typical multiphysics problem, such as the
terms in the \glspl{PDE}, the initial and boundary conditions, the material
properties, etc., are represented in \gls{MOOSE} as C++ objects. New objects
can inherit properties from related old objects to simplify implementation and
reduce code duplication. Overall, this approach
is helpful for many researchers, a significant fraction of whom do not possess
high-level programming backgrounds, as they are unencumbered by the
technical details and complexities involved in programing mesh handling
and \gls{PDE} solving in finite element analysis.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.5\textwidth]{central-core}
    \caption{2-D axisymmetric model of the MSFR. The red box indicates the
    central core region relevant to the modeling approach in Moltres.}
    \label{fig:core}
\end{figure}

@@

\begin{table}[htb!]
	\centering
	\caption{Neutron energy group upper bounds used in Serpent.}
	\begin{tabular}{c S}
		\toprule
		{Group number} & {Upper bound [MeV]}\\
		\midrule
		1 & 20\\
		2 & 2.2313\\
		3 & 0.4979\\
		4 & 0.0247875\\
		5 & 0.0055308\\
		6 & 0.0007485\\
		\bottomrule
	\end{tabular}
	\label{table:bound}
\end{table}

>>>>

This work demonstrates the \gls{MSFR} simulation capabilities of Moltres, a
multiphysics simulation tool for \glspl{MSR} \cite{lindsay_introduction_2018}.
To run simulations on Moltres, it requires input
group constant data from a neutron transport solver for the
multigroup neutron diffusion calculations, and a mesh file representing the
geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
for the former and Trelis \cite{noauthor_trelis_2018} for the latter.
This chapter provides brief introductions of
Serpent 2, MOOSE, and Moltres, and the general modeling approach for the
multiphysics simulations in Moltres.

@@

Serpent 2 \cite{leppanen_serpent_2014} is a continuous-energy Monte Carlo
neutron transport application under
active development led by the VTT Technical Research Centre of Finland. It was
created in 2004 for group constant generation in lattice geometries, and has
since grown to support more general capabilities. Serpent 2 is highly
parallelizable, supporting
both MPI and OpenMP parallel programming APIs. It has also been validated and
verified against experimental data and other well-established codes.

In Serpent 2, each neutron is tracked through a combination of
ray-tracing-based surface tracking and rejection sampling-based delta
tracking. Users may define the number of neutron histories and the number of
active and inactive cycles for each
simulation. Inactive cycles are required for fission source distribution
convergence, before interactions are tallied in the active cycles.
Interaction types and locations are
determined stochastically based on neutron interaction data from established
nuclear data libraries (e.g. ENDF \cite{chadwick_endf/b-vii.1_2011}, JEFF
\cite{oecd/nea_jeff-3.1.2_2014}). These nuclear data libraries provide
continuous-energy cross section data at discrete temperatures. Beyond the
discrete library temperatures, Serpent 2 has a built-in Doppler-broadening
preprocessor that extrapolates the relevant cross section data from a lower
temperature \cite{leppanen_serpent_2014}.

In the context of this work, Serpent 2 uses the JEFF-3.1.2 nuclear
data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
needed by Moltres. The relevant group constant data are collapsed into six
neutron energy groups, and calculated at discrete temperature values from 800
K to 1300 K at 100 K intervals. Table \ref{table:bound} shows the upper bounds
of each neutron energy group. The group constants relevant for neutronics
calculations in Moltres are the
macroscopic fission $\Sigma^f_{g}$, removal $\Sigma^r_{g}$, and scattering
$\Sigma^s_{g' \rightarrow g}$ neutron cross sections,
neutron diffusion coefficients $D_g$, average fission energies $\epsilon_g$,
average neutron yields $\nu$, inverse neutron speeds $1/v$, flux spectra
$\chi$, \gls{DNP} decay constants $\lambda_i$, and effective delayed neutron
fractions $\beta_{eff}$. These group constants are extracted from
the Serpent 2 output files using a Python script available from the Github
repository that holds the Moltres source code \cite{lindsay_moltres_2017}. The
script rewrites the group constants into a Moltres-compatible format.

Serpent 2 provides standard geometric surfaces (e.g. planes, cylinders, cones)
for defining reactor geometries. In this work, the reactor geometry is the
same axisymmetric \gls{MSFR} geometry from the paper by Fiorina et al.
\cite{fiorina_modelling_2014}, as shown in Fig. \ref{fig:msfrgeom}.

\begin{table}[htb!]
	\centering
	\caption{Neutron energy group upper bounds used in Serpent 2.}
	\begin{tabular}{c S}
		\toprule
		{Group number} & {Upper bound [MeV]}\\
		\midrule
		1 & 20\\
		2 & 2.2313\\
		3 & 0.4979\\
		4 & 0.0247875\\
		5 & 0.0055308\\
		6 & 0.0007485\\
		\bottomrule
	\end{tabular}
	\label{table:bound}
\end{table}

\section{MOOSE and Moltres}

Moltres \cite{lindsay_moltres_2017} is a \gls{MOOSE}-based
\cite{gaston_physics-based_2015} application for
coupled neutronics/thermal-hydraulics simulations of \glspl{MSR}. MOOSE
provides foundational framework for computationally solving \glspl{PDE} and
pre-built physics modules for common applications such as the Navier-Stokes
physics module. Moltres is also a MOOSE physics module for simulating
\glspl{MSR}. This section describes MOOSE, the Navier-Stokes physics module,
and Moltres.

\subsection{MOOSE}

\gls{MOOSE} \cite{gaston_physics-based_2015} is a highly parallelizable,
finite element framework developed at \gls{INL} for simplifying the process of
creating fully-coupled, non-linear, multiphysics solvers. The framework
provides a user-friendly interface for this task through object-oriented
programming in C++. All aspects of a typical multiphysics problem, such as the
terms in the \glspl{PDE}, the initial and boundary conditions, the material
properties, etc., are represented in \gls{MOOSE} as C++ objects. New objects
can inherit properties from related old objects to simplify implementation and
reduce code duplication. Overall, this approach
is helpful for many researchers, a significant fraction of whom do not possess
high-level programming backgrounds, as they are unencumbered by the
technical details and complexities involved in programing mesh handling
and \gls{PDE} solving in finite element analysis.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.5\textwidth]{central-core}
    \caption{2-D axisymmetric model of the MSFR. The red box indicates the
    central core region in the modeling approach in Moltres.}
    \label{fig:core}
\end{figure}

@@
